Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die Einhaltung der Echtzeit

### 2024-03-31 Masterarbeit_Robotik_Pamuk.tex
Application Context and  Produktkatalog

\bigskip \noindent In these RTOS systems, task scheduling is based on so-called priority-based preemptive scheduling. Each task in a software application is assigned a priority. A higher priority means that a faster response is required. Pre-emptive task scheduling ensures a very fast response. Preemptive means that the scheduler can stop a currently running task at any point if it recognises that another task needs to be executed immediately. The basic rule on which priority-based preemptive scheduling is based is that the task with the highest priority that is ready to run is always the task that must be executed. So if both a task with a lower priority and a task with a higher priority are ready to run, the scheduler ensures that the task with the higher priority runs first. The lower priority task is only executed once the higher priority task has been processed. Real-time systems are usually categorised as either soft or hard real-time systems. The difference lies exclusively in the consequences of a violation of the time limits.

\bigskip \noindent Hard real-time is when the system stops operating if a deadline is missed, which can have catastrophic consequences. These consequences include endangering people, damaging machinery, impacting the health and integrity of the environment. Typical examples of this are some control systems in aeroplanes or cars, e.g. internal combustion engines.

\bigskip \noindent Soft real-time exists when a system continues to function even if it cannot perform the tasks within a specified time. If the system has missed the deadline, this has no critical consequences. The system continues to run, although it does so with undesirably lower output quality. Examples of this are multimedia systems where the occasional deviation from a playback rate of 25 frames/sec. leads to stuttering or similar.

%In der heutigen industriellen Fertigung und Automatisierung sind Robotersysteme sehr etabliert und von entscheidender Bedeutung. Roboter müssen auf ihre Umgebung reagieren und zeitkritische Aufgaben innerhalb strenger Zeitvorgaben ausführen. Verzögerungen oder Fehler können in einigen Fällen katastrophale Konseqenzen haben. Traditionelle Betriebssysteme, wie Windows oder Linux sind für diese Art von Echtzeitanforderungen oft nicht geeignet, da sie keine deterministischen Ausführungszeiten garantieren können. Daher sind Echtzeit-Betriebssysteme nötig, die speziell dafür konzipiert sind, auf Ereignisse innerhalb fester Zeitschranken zu reagieren und Prozesse mit hoher Priorität bevorzugt auszuführen. Bei diesen RTOS-Systemen basiert das Task-Scheduling auf dem sogenannten prioritätsbasierten präemptiven Scheduling. Jeder Task in einer Software-Applikation ist eine Priorität zuzuweisen. Eine höhere Priorität bedeutet, dass eine schnellere Reaktion erforderlich ist. Durch das präemptive Task-Scheduling wird eine sehr schnelle Reaktion sichergestellt. Präemptiv bedeutet, dass der Scheduler eine gerade laufende Task an jeder Stelle anhalten kann, wenn er erkennt, dass eine andere Task sofort ausgeführt werden muss. Die Grundregel, auf der das prioritätsbasierte präemptive Scheduling basiert, besagt, dass die ausführbereite Task mit der höchsten Priorität immer die Task ist, die ausgeführt werden muss. Wenn also sowohl eine Task mit niedriger als auch eine Task mit höherer Priorität ausführbereit sind, sorgt der Scheduler dafür, dass zuerst die Task mit der höheren Priorität läuft. Die Task mit niedrigerer Priorität wird erst dann ausgeführt, wenn die höherpriore Task verarbeitet wurde.

%\noindent Echtzeitsysteme werden in der Regel entweder als weiches oder als hartes Echtzeitsystem kategorisiert. Der Unterschied liegt ausschließlich in den Folgen einer Verletzung der Zeitschranken.

%\noindent Harte Echtzeit liegt vor, wenn das System seinen Betrieb einstellt, wenn ein Termin verpasst wurde, was katastrophale Folgen haben kann. Zu diesen Folgen zählen die Gefährdung von Menschen, die Beschädigung von Maschinen, also Auswirkungen auf Gesundheit und Unversehrtheit der Umgebung. Typische Beispiele hierfür sind einige Steuerungssysteme im Flugzeug oder im Auto, z.B. bei der Verbrennungsmaschine.

%\noindent Weiche Echtzeit besteht, wenn ein System weiter funktionsbereit ist, selbst wenn es die Aufgaben innerhalb einer festgelegten Zeit nicht ausführen kann. Wenn das System den Termin verpasst hat, hat dies keine kritischen Konsequenzen. Das System läuft weiter, obwohl es dies mit unerwünscht geringerer Ausgabenqualität tut. Beispiele hierfür sind Multimediasysteme, bei denen das gelegentlich Abweichen von einer Abspielrate von 25 Bildern/sek. zu einem Ruckeln o.ä. führt.

\clearpage
\section{Application Context}
This master's thesis was written at SIGMATEK GmbH \& Co KG \cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}.

@@

\clearpage 
\chapter{Salamander Operating System}

@@

\begin{table}[ht]
    \centering
    \caption{Overview of the priority groups and their relationships}
    \label{tab:priorities}
    \begin{tabular}{|c|c|}
        \hline
        \textbf{Prioritätsgruppe} & \textbf{Bereich}  \\ \hline
        Xenomai Priorität & 0 bis 99  \\ \hline
        Linux RT Priorität & 1 bis 99 \\ \hline
        Linux (Nice Level) Priorität & -20 bis 19  \\ \hline
        RTK Priorität & 0 bis 14 \\ \hline
    \end{tabular}

\end{table}

@@

\begin{table}[H]
  \centering
  \begin{tabular}{|l|p{0.62\textwidth}|}
  \hline
  \textbf{Exit Reason} & \textbf{Description}\\
  \hline
  APIC\_WRITE & Triggered when the guest writes to its APIC. \\ \hline
  EXTERNAL\_INTERRUPT & Triggered by external hardware interrupts. \\ \hline
  HLT & Triggered when the guest executes the HLT instruction. \\ \hline
  EPT\_MISCONFIG & Triggered by a misconfiguration in the EPT. \\ \hline
  PREEMPTION\_TIMER & Triggered when the host's preemption timer expires. \\ \hline
  PAUSE\_INSTRUCTION & Triggered when the PAUSE instruction is executed. \\ \hline
  EPT\_VIOLATION & Triggered by a violation of the EPT permission settings. \\ \hline
  IO\_INSTRUCTION & Triggered when the guest executes an I/O instruction. \\ \hline
  EOI\_INDUCED & Triggered when an EOI signal is sent to the APIC. \\ \hline
  MSR\_READ & Triggered when the guest reads from a MSR. \\ \hline
  CPUID & Triggered when the guest executes the CPUID instruction.  \\ \hline
  \end{tabular}
  \caption{Description of kvm\_exit reasons}
  \label{tab:kvm_exit}
  \end{table}

>>>>

\bigskip \noindent In these RTOS systems, task scheduling is based on so-called priority-based preemptive scheduling. Each task in a software application is assigned a priority. A higher priority means that a faster response is required. Preemptive task scheduling ensures a very fast response. Preemptive means that the scheduler can stop a currently running task at any point if it recognizes that another task needs to be executed immediately. The basic rule on which priority-based preemptive scheduling is based is that the task with the highest priority that is ready to run is always the task that must be executed. So if both a task with a lower priority and a task with a higher priority are ready to run, the scheduler ensures that the task with the higher priority runs first. The lower priority task is only executed once the higher priority task has been processed. Real-time systems are usually categorized as either soft or hard real-time systems. The difference lies exclusively in the consequences of a violation of the time limits.

\bigskip \noindent Hard real-time is when the system stops operating if a deadline is missed, which can have catastrophic consequences. Soft real-time exists when a system continues to function even if it cannot perform the tasks within a specified time. If the system has missed the deadline, this has no critical consequences. The system continues to run, although it does so with undesirably lower output quality.

\clearpage
\section{Application Context}
This master's thesis was written at SIGMATEK GmbH \& Co KG \cite{pixelartSIGMATEKKompletteAutomatisierungssysteme}. This section provides necessary information about the starting point and boundary conditions of the thesis, as the work should be reproducible with the means used.

SIGMATEK uses its own customized Linux distribution to be run on their self-manufactured CPUs, namely Salamander 4. This operating system

, to employ hard real-time in their Salamander4 is virtualized through third party services.

The details of this operating system are explained in section \ref{sec:salamander4}

@@

\clearpage
\chapter{Salamander Operating System}\label{sec:salamander4}

@@

\begin{table}[ht]
	\centering
	\caption{Overview of the priority groups and their relationships}
	\label{tab:priorities}
	\begin{tabular}{|c|c|}
		\hline
		\textbf{Prioritätsgruppe}    & \textbf{Bereich} \\ \hline
		Xenomai Priorität            & 0 bis 99         \\ \hline
		Linux RT Priorität           & 1 bis 99         \\ \hline
		Linux (Nice Level) Priorität & -20 bis 19       \\ \hline
		RTK Priorität                & 0 bis 14         \\ \hline
	\end{tabular}
\end{table}

@@

\begin{table}[H]
	\centering
	\begin{tabular}{|l|p{0.62\textwidth}|}
		\hline
		\textbf{Exit Reason} & \textbf{Description}                                     \\
		\hline
		APIC\_WRITE          & Triggered when the guest writes to its APIC.             \\ \hline
		EXTERNAL\_INTERRUPT  & Triggered by external hardware interrupts.               \\ \hline
		HLT                  & Triggered when the guest executes the HLT instruction.   \\ \hline
		EPT\_MISCONFIG       & Triggered by a misconfiguration in the EPT.              \\ \hline
		PREEMPTION\_TIMER    & Triggered when the host's preemption timer expires.      \\ \hline
		PAUSE\_INSTRUCTION   & Triggered when the PAUSE instruction is executed.        \\ \hline
		EPT\_VIOLATION       & Triggered by a violation of the EPT permission settings. \\ \hline
		IO\_INSTRUCTION      & Triggered when the guest executes an I/O instruction.    \\ \hline
		EOI\_INDUCED         & Triggered when an EOI signal is sent to the APIC.        \\ \hline
		MSR\_READ            & Triggered when the guest reads from a MSR.               \\ \hline
		CPUID                & Triggered when the guest executes the CPUID instruction. \\ \hline
	\end{tabular}
	\caption{Description of kvm\_exit reasons}
	\label{tab:kvm_exit}
\end{table}
